---
title: "Between-dataset comparisons"
output: 
  github_document
params:
  on_hpg: [FALSE]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

if(params$on_hpg) {

all_di <- list() 
datasets <- c("bbs", "fia_short", "gentry", "mcdb", "misc_abund_short", "portal_plants")

#datasets <- c("fia", "gentry", "mcdb", "misc_abund_short", "portal_plants")

for(i in 1:length(datasets)) {
  
  this_dataset = datasets[i]
  if(this_dataset == "misc_abund_short") {
    cache_loc = "miscabund"
  } else if(this_dataset == "fia_short") {
    
    cache_loc = "fia"
  } else {
    cache_loc = this_dataset
  }
  
  ## Set up the cache and config
  db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite")))
  cache <- storr::storr_dbi("datatable", "keystable", db)
  
    all_di[[i]] <- readd(all_di_obs, cache = cache)
  
  all_di[[i]] <- filter(all_di[[i]], source == "observed")
  
  this_dat <- readd(paste0("dat_s_dat_", this_dataset), cache = cache, character_only = T)
  sv <- get_statevars(this_dat)
  all_di[[i]] <- left_join(all_di[[i]], sv, by = c("sim", "site", "source", "singletons", "dat"))
  
  
  DBI::dbDisconnect(db)
  rm(cache)
  rm(db)
}


all_di <- bind_rows(all_di)

write.csv(all_di, "all_di.csv", row.names = F)
} else {
  all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
}
```

```{r histograms, fig.height = 6, fig.width =6}

skewhists <- ggplot(data = all_di, aes(x = skew_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500)) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Skewness")
skewhists


simpsonhists <- ggplot(data = all_di, aes(x = simpson_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500)) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Simpson")
simpsonhists

```
 
Observations from these histograms:

* FIA really struggles to get 2500 samples
* Portal is the *most* concentrated of all the datasets.
* Then MCDB and misc, then gentry, then bbs. FIA is hard to place but looks idiosyncratic.
* Curious about how variation in %ile maps on to S and N, *particularly* for Gentry and BBS. 
* Skewness is less extreme than Simpson (more density in the in-between). 
* Singletons doesn't appear to be terrifically impactful. From now on filtering to *unaltered* vectors.

```{r s and n mapping, fig.width = 12, fig.height = 10}

all_di <- filter(all_di, !singletons)

skew_s_n_maps <- ggplot(data = all_di, aes(x = s0, y = n0, color = skew_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free") + 
  ggtitle("Skewness") +
  scale_color_viridis_c(option = "plasma", end= .75)

skew_s_n_maps


simpson_s_n_maps <- ggplot(data = all_di, aes(x = s0, y = n0, color = simpson_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free") + 
  ggtitle("Simpson") +
  scale_color_viridis_c(option = "plasma", end= .75, direction = -1)

simpson_s_n_maps

```

The less extreme (low skew, high simpson) vectors appear vaguely collected in the lower right for Gentry and BBS. Those are the regions with relatively high N/S, aka low average abundance, aka a relatively small feasible set. 


Conversely, for MCDB and Misc, it almost looks like the very least skewed elements are way down at tiny abundances?

Do percentiles scan with S/N?

```{r avg abund}

all_di <- mutate(all_di, avg_abund = n0/s0)

skew_avgabund <- ggplot(data = all_di, aes(x = avg_abund, y = skew_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  ggtitle("Skew") +
  facet_wrap(vars(dat), scales = "free")

skew_avgabund


simpson_avgabund <- ggplot(data = all_di, aes(x = avg_abund, y = simpson_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  ggtitle("Simpson") +
  facet_wrap(vars(dat), scales = "free")

simpson_avgabund

```
There is perhaps a constraint, but that might be because you rarely get v high average abundances. You seem to rarely get low skewness or high Simpson at high average abundance, which scans with the heatmaps. 

Do Simpson and skewness correspond?
```{r do skew and simp correlate}

ss_plot <- ggplot(data = all_di, aes(x= skew_percentile, y = simpson_percentile, color = dat)) +
  geom_point(alpha = .1) +
  scale_color_viridis_d(end = .8) +
  theme_bw() +
  ggtitle("Skew v Simp")

ss_plot

ss_facet_plot <- ss_plot +
  facet_wrap(vars(dat))

ss_facet_plot
```

There's a weak relationship between Simpson and skewness, but a lot of noise. They are not substitutable. 

Why is Gentry so strangely bimodal? It's like a U, when all the others tend towards one end or the other. 

Why are BBS and mammals less frequently squished than Portal?


Looking at range of skewness and Simpsons:

```{r ranges}

skew_range_hist <- ggplot(data = filter(all_di, nsamples > 2000), aes(x = skew_range, y = skew_percentile, color = dat)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "lm", se =F) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(dat), scales = "free") + 
  ggtitle("Skewness")

skew_range_hist



simpson_range_hist <- ggplot(data = filter(all_di, nsamples > 2000), aes(x = simpson_range, y = simpson_percentile, color = dat)) +
  geom_point(alpha = .2) +
    geom_smooth(method = "lm", se =F) +

  theme_bw() +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(dat), scales = "free") +
  ggtitle("Simpson")

simpson_range_hist

```

I added the ranges because I thought the percentiles might be constrained somewhat by the range of values represented in the feasible set. It doesn't look to me like there is a strong relationship between range and %ile. 
