---
title: "Dissimilarity of FS vs. actual"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    # github_document:
    #    df_print: kable
    #    toc: true
    word_document:
        df_print: kable
        toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F)

all_ct <-  read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"), stringsAsFactors = F)


#fia_ct <- read.csv(here::here("fia_cts.csv"))

#all_ct <- rbind(all_ct, fia_ct)
all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  left_join(all_ct) 

all_di <- all_di %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl),
         skew_percentile_mean = mean(skew_percentile, skew_percentile_excl),
         simpson_percentile_mean = mean(simpson_percentile,simpson_percentile_excl),
         shannon_percentile_mean = mean(shannon_percentile, shannon_percentile_excl),
         nsingletons_percentile_mean = mean(nsingletons_percentile, nsingletons_percentile_excl),) %>%
  ungroup()

all_di <- all_di %>%
  mutate(in_fia = ifelse(Dataset == "FIA", "FIA", "Other datasets"))

```


```{r}
plot_percentile_hist <- function(di_df, col_name, plot_name, tails = 2, facetvar = "Dataset", min_s0 = 0) {

  if(tails == 2) {
    cutoff_percentiles = c(2.5, 97.5)
    min_nparts = 40
  } else if (tails== 1) {
    cutoff_percentiles = c(95)
    min_nparts = 20
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[col_name]]),
           facetvar = di_df[[facetvar]])


  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    xlab("") +
    ylab("") +
    geom_vline(xintercept = cutoff_percentiles, color = "red") +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+
    theme(plot.title = element_text(size=10))



}

all_di <- all_di %>%
  mutate(`Observed \npercentile score` = ifelse(real_po_percentile_excl > 95, "High (> 95)", "Less than 95"))

```

```{r, fig.dim = c(7,7)}
ggplot(all_di, aes(sim_pos_from_best, real_po, color = `Observed \npercentile score`)) +
  geom_point(alpha = .3) +
  geom_point(data = filter(all_di, s0 < 0)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_color_viridis_d(end = .8, direction = -1) +
  xlab("Mean of elements from feasible set") +
  ylab("Observed value") +
  ggtitle("Dissimilarity to the central tendency") +  theme(legend.position = "bottom")+
  facet_wrap(vars(Dataset))
#,
#
#
# ggplot(filter(all_di, nparts < 10 ^ 50), aes(nparts, sim_pos_from_best, color = Dataset)) +
#   geom_point(alpha = .3) +
#   geom_point(data = filter(all_di, s0 < 0)) +
#   scale_color_viridis_d(end = .9) +
#   xlab("Number of elements in the feasible set") +
#   ylab("Mean dissimilarity of feasible set to central tendency") +
#   scale_x_log10() +
#   ggtitle("Over the size of the feasible set")
# ), ncol = 1,
#top = textGrob("Dissimilarity to central tendency"))

```

**Figure S4**. Observed SADs are often much more dissimilar to the central tendency of their feasible sets (y-axis) than the mean dissimilarity of samples from the feasible set and the central tendency (x-axis). Dissimilarity can range from 0-1. The black line is the 1:1 line. Of observed SADs that are more disssimilar to the central tendency than are 95% of samples from the feasible set (green points), the absolute increase in dissimlarity ranges from .04 to .6. These observed SADs are from 1.4 to 9.7 times more dissimilar to the central tendency than the maen of samples from the feasible set. 
