---
title: "Unusual Gentry communities"
#author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    # github_document:
    #    df_print: kable
    #    toc: true
   word_document:
        df_print: kable
        toc: false
        reference_docx: style_reference.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())

all_di <- read.csv(here::here("analysis", "reports", "submission2", "all_di.csv"), stringsAsFactors = F)

all_ct <-  read.csv(here::here("analysis","reports", "submission2", "all_ct.csv"), stringsAsFactors = F)


#fia_ct <- read.csv(here::here("fia_cts.csv"))

#all_ct <- rbind(all_ct, fia_ct)
all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  left_join(all_ct) 

all_di <- all_di %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl),
         skew_percentile_mean = mean(skew_percentile, skew_percentile_excl),
         simpson_percentile_mean = mean(simpson_percentile,simpson_percentile_excl),
         shannon_percentile_mean = mean(shannon_percentile, shannon_percentile_excl),
         nsingletons_percentile_mean = mean(nsingletons_percentile, nsingletons_percentile_excl),) %>%
  ungroup()

all_di <- all_di %>%
  mutate(in_fia = ifelse(Dataset == "FIA", "FIA", "Other datasets"))

```


```{r, fig.dim = c(3, 6)}
plot_percentile_hist <- function(di_df, col_name, plot_name, tails = 2, facetvar = "Dataset", min_s0 = 0) {

  if(tails == 2) {
    cutoff_percentiles = c(2.5, 97.5)
    min_nparts = 40
  } else if (tails== 1) {
    cutoff_percentiles = c(95)
    min_nparts = 20
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[col_name]]),
           facetvar = di_df[[facetvar]])


  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    xlab("") +
    ylab("") +
    geom_vline(xintercept = cutoff_percentiles, color = "red") +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+
    theme(plot.title = element_text(size=10))



}

all_di <- all_di %>%
  mutate(`Observed \npercentile score` = ifelse(real_po_percentile_excl > 95, "High (> 95)", "Less than 95"))

```

```{r, fig.dim = c(3, 6)}

plot_narrowness <- function(di_df, yvar, yvar_name) {

  if(grepl("sim_pos", yvar)) {
    min_nparts = 20
    ylabel = "Mean dissimilarity"
  } else {
    min_nparts = 40
    ylabel = "Breadth index"
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[yvar]])) %>%
    filter(nparts > min_nparts,
           nparts < 10 ^ 50)


  ggplot(di_df, aes(nparts, response, color = Dataset)) +
    geom_point(data = filter(di_df, dat == "fia"), alpha = .1) +
    geom_point(data = filter(di_df, dat != "fia"), alpha = .1) +
    geom_point(data = filter(di_df, s0 < 0)) +
    scale_color_viridis_d(end = .9) +
    xlab("") +
    ylab(ylabel) +
    scale_x_log10() +
    ggtitle(yvar_name) +
    theme(legend.position = "none")+
    theme(plot.title = element_text(size=10))
}

plot_narrowness_legend <- function(di_df) {

  legend_df <- di_df %>%
    select(Dataset) %>%
    distinct() %>%
    mutate(ymark = dplyr::row_number())

  ggplot(legend_df, aes(1, ymark, color = Dataset)) +
    geom_label(aes(x = 2, y = ymark, label = Dataset)) +
    geom_point(size = 4) +
    scale_color_viridis_d(end = .9) +
    xlab("") +
    ylab("") +
    theme(legend.position = "none") +
    xlim(1, 3) + theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    ) +
    ggtitle("Dataset")

}
#
# gridExtra::grid.arrange(grobs = list(
#   plot_narrowness(all_di, "sim_pos_from_best", "Dissimilarity to central tendency"),
#   plot_narrowness_legend(all_di)),
#   ncol = 1)
```

```{r}

plot_narrowness_hist <- function(di_df, col_name, plot_name, facetvar = "Dataset", min_s0 = 0) {

  if(grepl("sim_pos", col_name)) {
    xvarname = "Mean dissimilarity"
    min_nparts = 40
  } else {
    min_nparts = 20
    xvarname = "Breadth index"
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[col_name]]),
           facetvar = di_df[[facetvar]])


  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    ylab("") +
    xlab(xvarname) +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+
    theme(plot.title = element_text(size=10))

}

```



The Gentry communities are extreme relative to the other datasets (in grey), in that they have high species richness and low abundance. Of these, the most extreme communities have extremely low average abundance (e.g. N/S < 3).

```{r, fig.dim = c(4,4)}

gentry_di <- filter(all_di, Dataset == "Gentry")

gentry_di <- gentry_di %>%
  mutate(low_avg_abund = ifelse(n0/s0 < 3, "N:S < 3", "N:S > 3")) %>%
  mutate(`Low average \nabundance` = low_avg_abund)

ggplot(gentry_di, aes(s0, n0, color = `Low average \nabundance`)) +
  geom_point(data = all_di, inherit.aes = F, aes(s0, n0), color = "grey", alpha = .1) +  geom_point() +
scale_x_log10() + scale_y_log10() + xlab("Species richness (S); note log scale") + ylab("Total abundance (N); note log scale") + ggtitle("Gentry communities", subtitle = "(in color)") + theme(legend.position = "bottom") + scale_color_viridis_d(end = .8)

```

This figure shows all of the communities across all our datasets in terms of S and N. Gentry communities are in color, color coded by whether they have low average abundance (N/S < 3). 


These low-average-abundance communities - the dark purple dots in the plot above - account for most of the signal of extreme values in the unusual direction for various metrics - that is, a **low** number of rare species and **low** skewness, and **high** Shannon and Simpson evenness. Gentry communities with average abundance > 3 tend to show the signal in the same direction as most other communities. 


```{r, fig.dim = c(7,7)}
fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_percentile_hist(gentry_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "low_avg_abund"),
  plot_percentile_hist(gentry_di, "skew_percentile_mean", "\nSkewness", facetvar = "low_avg_abund", min_s0 = 3),
  plot_percentile_hist(gentry_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "low_avg_abund"),
  plot_percentile_hist(gentry_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "low_avg_abund")
), ncol = 2,
 left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))



```


These low-average abundance communities have unusual statistical baselines. For example, samples from these feasible sets have a very high proportion of rare species:

```{r, fig.dim = c(4.5,4)}


all_di <- all_di %>%
  mutate(`Average proportion of rare species \nin the feasible set` = nsingletons_mean / s0)

ggplot(all_di, aes(s0, n0, color = `Average proportion of rare species \nin the feasible set`)) +
  geom_point(alpha = .4) +
scale_x_log10() + 
  scale_y_log10() + 
  scale_color_viridis_c(option = "plasma", end = .8) +
  xlab("Species richness (S); note log scale") + ylab("Total abundance (N); note log scale") + 
  theme(legend.position = "bottom", plot.margin =unit(c(4,4,4,4), units = "mm"))

```
